---
title: "Capstone Project"
author: "Ji-Eun Park"
date: "5/17/2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(irlba)
library(splatter)
library(scater)
library(clusterSim)
library(SingleCellExperiment)
library(clusterExperiment)
library(readr)
```


```{r}
dat2 <- read_rds('/Users/jieun/Desktop/CapstoneProject/datasets/kolodziejczyk.rds')
dat2 <- dat2@assays$data$counts
dat2 <- as.matrix(dat2)
dat2 <- na.omit(dat2)
dat2 <- dat2[rowSums(dat2)>0,]
```

## Parameter Estimate
```{r}
params_2 <- splatEstimate(dat2) # use the data to generate parameters
pc <- c(2,5,10,20,50) # Try various numbers for PC = (2,5,10,20,50)
simnum <- 3 # number of simulations
ks <- 2:20
```

## Simulation
```{r}
sim_groups_2 <- lapply(1:simnum, function(i) splatSimulate(params_2,group.prob = c(0.3,0.3,0.4), method = "groups",verbose = FALSE,de.facLoc=.25, de.facScale=1,seed=i)) #3 simulated data

sim_norm_2 <- lapply(sim_groups_2,normalise) #normalise simulated data

filtered_counts_2 <- lapply(sim_norm_2, function(sim) sim[rowSums(logcounts(sim))>0,]) #sort out rows that are not all 0 logcounts

sim_SCE_2 <- lapply(filtered_counts_2, function(sim) runPCA(sim, ncomponents = 50, method = "irlba", ntop = 1000)) # compute PCA

# (Non-sequential) clustering with kmeans and PAM with k from 2 to 20.
# Take only number of components of interest (2,5,10,20,50)
clust_res_2 <- lapply(sim_SCE_2, function(sim) clusterMany(sim,ks=2:20,alphas=0.1,betas=0.8,minSizes=1,
            clusterFunction=c("kmeans","pam"), sequential=FALSE,subsample=FALSE,
                  reduceMethod="PCA",nReducedDims=c(2,5,10,20,50),verbose=TRUE)) 


# (Sequential) clustering with kmeans and PAM with k from 2 to 20.
# Take only number of components of interest (2,5,10,20,50)
clust_res_seq_2 <- lapply(sim_SCE_2, function(sim) clusterMany(sim,ks=2:20,alphas=0.1,betas=0.8,minSizes=1,
            clusterFunction=c("kmeans","pam"), sequential=TRUE,subsample=FALSE,
                  reduceMethod="PCA",nReducedDims=c(2,5,10,20,50),verbose=TRUE))

# With ClusterExperiment object generate cluster matrix showing the clustered group names of each cell(row)
clust_res_mat_2 <- lapply(clust_res_2,function(sim) clusterMatrix(sim))
clust_res_seq_mat_2 <- lapply(clust_res_seq_2,function(sim) clusterMatrix(sim))

# For sequential clustering methods, generate an list of empty matrices   
clust_res_mat_seq_NA_2 <- list("1"=matrix(nrow=nrow(clust_res_seq_mat_2[[1]]),ncol=ncol(clust_res_seq_mat_2[[1]])),"2"=matrix(nrow=nrow(clust_res_seq_mat_2[[2]]),ncol=ncol(clust_res_seq_mat_2[[2]])),"3"=matrix(nrow=nrow(clust_res_seq_mat_2[[3]]),ncol=ncol(clust_res_seq_mat_2[[3]])))

# Change group -1 to NA
for (a in 1:simnum) {
  for (b in 1:nrow(clust_res_seq_mat_2[[1]])) {
    for (c in 1:ncol(clust_res_seq_mat_2[[1]])) {
      clust_res_mat_seq_NA_2[[a]][b,c] <- ifelse(clust_res_seq_mat_2[[a]][b,c]=="-1",NA,clust_res_seq_mat_2[[a]][b,c])
    }
  }
}

# assign colnames to new list of matrices
for (i in 1:simnum) {
  colnames(clust_res_mat_seq_NA_2[[i]]) <- colnames(clust_res_seq_mat_2[[i]])
}

# if a column is all NA change everything is Group1
clust_res_mat_NA_fixed_2 <- lapply(clust_res_mat_seq_NA_2, function(x) {
  idx <- which(colSums(is.na(x)) == NROW(x))
  x[,idx] <- 1
  return(x)
})

# divide the matrices in list to kmeans matrices and pam matrices
kmenasMat_2 <- lapply(clust_res_mat_2, function(sim) sim[,1:95])
pamMat_2 <- lapply(clust_res_mat_2, function(sim) sim[,96:190])

kmeansMat2_2 <- lapply(clust_res_mat_NA_fixed_2, function(sim) sim[,1:95]) 
pamMat2_2 <- lapply(clust_res_mat_NA_fixed_2, function(sim) sim[,96:190])

ks=2:20
```

## Clustering
### Non-sequential Kmeans
```{r}
# kmeans for each pc regardless of k
kmeans_by_pc_2 <- lapply(1:length(pc), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(kmenasMat_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

avg_kmean_by_pc_2 <- lapply(1:length(pc), function(i) {
    lapply(1:simnum,function (j) {
      mean(unlist(kmeans_by_pc_2[[i]][[j]]))
    })
})

avg_kmean_by_pc_2 <- lapply(1:length(pc), function(i){
  mean(unlist(avg_kmean_by_pc_2[[i]]))
})

avg_kmean_by_pc_2 <- unlist(avg_kmean_by_pc_2)
```

### Non-sequential PAM
```{r}
# pam for each pc regardless of k
pam_by_pc_2 <- lapply(1:length(pc), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(pamMat_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

avg_pam_by_pc_2 <- lapply(1:length(pc), function(i) {
    lapply(1:simnum,function (j) {
      mean(unlist(pam_by_pc_2[[i]][[j]]))
    })
})

avg_pam_by_pc_2 <- lapply(1:length(pc), function(i){
  mean(unlist(avg_pam_by_pc_2[[i]]))
})

# Unlist for plotting
avg_pam_by_pc_2 <- unlist(avg_pam_by_pc_2)
```

### Sequantial Kmeans
```{r}
# kmeans for each pc regardless of k
kmeans_by_pc2_2 <- lapply(1:length(pc), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(kmeansMat2_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

avg_kmean_by_pc2_2 <- lapply(1:length(pc), function(i) {
    lapply(1:simnum,function (j) {
      mean(unlist(kmeans_by_pc2_2[[i]][[j]]))
    })
})

avg_kmean_by_pc2_2 <- lapply(1:length(pc), function(i){
  mean(unlist(avg_kmean_by_pc2_2[[i]]))
})

avg_kmean_by_pc2_2 <- unlist(avg_kmean_by_pc2_2)
```

### Sequential PAM
```{r}
pam_by_pc2_2 <- lapply(1:length(pc), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(pamMat2_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

avg_pam_by_pc2_2 <- lapply(1:length(pc), function(i) {
    lapply(1:simnum,function (j) {
      mean(unlist(pam_by_pc2_2[[i]][[j]]))
    })
})

avg_pam_by_pc2_2 <- lapply(1:length(pc), function(i){
  mean(unlist(avg_pam_by_pc2_2[[i]]))
})

# Unlist for plotting
avg_pam_by_pc2_2 <- unlist(avg_pam_by_pc2_2)
```

## Plots for Average Rand Index(Y) for each number of Component(X) 
### Non-Sequential
```{r}
# Make a dataframe including all Indexes
avg_ari_2 <- data.frame(c(rep("kmeans",length(pc)),rep("pam",length(pc))),rep(pc,2),c(avg_kmean_by_pc_2,avg_pam_by_pc_2))

colnames(avg_ari_2) <- c("method","pc","index")

# Plot for AARI
ggplot(avg_ari_2,aes(x=pc,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

### Sequential
```{r}
# Make a dataframe including all Indexes
avg_ari2_2 <- data.frame(c(rep("kmeans",length(pc)),rep("pam",length(pc))),rep(pc,2),c(avg_kmean_by_pc2_2,avg_pam_by_pc2_2))

colnames(avg_ari2_2) <- c("method","pc","index")

# Plot for AARI
ggplot(avg_ari2_2,aes(x=pc,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

## Clustering result (for each k, PC fixed to 5)
### Non-sequential Kmeans
```{r}
## With PC=5
# kmeans by regardless of pc

kmeans_by_k3_2 <- lapply(1:length(ks), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(kmenasMat_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

# Get the average for each PCs
avg_kmeans_by_k3_2 <- lapply(1:length(ks), function(i) {
    lapply(1:simnum,function (j) {
  mean(unlist(kmeans_by_k3_2[[i]][[j]]))
    })
})

avg_kmeans_by_k3_2 <- lapply(1:length(ks), function(i) {
  mean(unlist(avg_kmeans_by_k3_2[[i]]))
})

avg_kmeans_by_k3_2 <- unlist(avg_kmeans_by_k3_2)
```

### Non-sequential PAM
```{r}
# kmeans regardless of pc
pam_by_k3_2 <- lapply(1:length(ks), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(pamMat_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

# Get the average for each PCs
avg_pam_by_k3_2 <- lapply(1:length(ks), function(i) {
  lapply(1:length(simnum), function (j){
      mean(unlist(pam_by_k3_2[[i]][[j]]))
  })
})

avg_pam_by_k3_2 <- lapply(1:length(ks), function(i) {
      mean(unlist(avg_pam_by_k3_2[[i]]))
})

# Unlist for plotting
avg_pam_by_k3_2 <- unlist(avg_pam_by_k3_2)
```

### Sequential Kmeans
```{r}
kmeans_by_k4_2 <- lapply(1:length(ks), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(kmeansMat2_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

# Get the average for each PCs
avg_kmeans_by_k4_2 <- lapply(1:length(ks), function(i) {
    lapply(1:simnum,function (j) {
  mean(unlist(kmeans_by_k4_2[[i]][[j]]))
    })
})

avg_kmeans_by_k4_2 <- lapply(1:length(ks), function(i) {
  mean(unlist(avg_kmeans_by_k4_2[[i]]))
})

avg_kmeans_by_k4_2 <- unlist(avg_kmeans_by_k4_2)
```

### Sequential PAM
```{r}
pam_by_k4_2 <- lapply(1:length(ks), function (i) {
  lapply(1:simnum,function (j) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(pamMat2_2[[j]][,col_num],sim_norm_2[[j]]$Group)
    })
  })
})

# Get the average for each PCs
avg_pam_by_k4_2 <- lapply(1:length(ks), function(i) {
  lapply(1:length(simnum), function (j){
      mean(unlist(pam_by_k4_2[[i]][[j]]))
  })
})

avg_pam_by_k4_2 <- lapply(1:length(ks), function(i) {
      mean(unlist(avg_pam_by_k4_2[[i]]))
})

# Unlist for plotting
avg_pam_by_k4_2 <- unlist(avg_pam_by_k4_2)
```

## Plots for Average Rand Index(Y) for each k (X)
### Non-Sequential
```{r}
# Make a dataframe including all Indexes
avg_ari3_2 <- data.frame(c(rep("kmeans",length(ks)),rep("pam",length(ks))),rep(ks,2),c(avg_kmeans_by_k3_2,avg_pam_by_k3_2))

colnames(avg_ari3_2) <- c("method","K","index")

# Plot for AARI
ggplot(avg_ari3_2,aes(x=K,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

### Sequential
```{r}
avg_ari4_2 <- data.frame(c(rep("kmeans",length(ks)),rep("pam",length(ks))),rep(ks,2),c(avg_kmeans_by_k4_2,avg_pam_by_k4_2))

colnames(avg_ari4_2) <- c("method","K","index")

# Plot for AARI
ggplot(avg_ari4_2,aes(x=K,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

```{r include=FALSE}
save(avg_ari_2,avg_ari2_2,avg_ari3_2,avg_ari4_2, file="/Users/jieun/Desktop/CapstoneProject/FinalScripts/data2results.Rdata")
```
 