---
title: "Capstone Project"
author: "Ji-Eun Park"
date: "5/17/2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(irlba)
library(splatter)
library(scater)
library(clusterSim)
library(SingleCellExperiment)
library(clusterExperiment)
library(readr)
```


### 3. Dataset3

```{r}

real_groups3 <- read_rds('/Users/jieun/Desktop/CapstoneProject/datasets/patel.rds')

# log count data with filters ( logcounts greater than 0 )
filtered_counts3 <- real_groups3[rowSums(logcounts(real_groups3))>0,]

sim_SCE_3 <- runPCA(filtered_counts3, ncomponents = 50, method = "irlba", ntop = 1000)
pc = c(2,5,10,20,50)
ks = 2:20

```

```{r}
# (Non-sequential) clustering with kmeans and PAM with k from 2 to 20.
# Take only number of components of interest (2,5,10,20,50)
clust_res_3 <-clusterMany(sim_SCE_3,ks=2:20,alphas=0.1,betas=0.8,minSizes=1,
            clusterFunction=c("kmeans","pam"), sequential=FALSE,subsample=FALSE,
                  reduceMethod="PCA",nReducedDims=c(2,5,10,20,50),verbose=TRUE) 


# (Sequential) clustering with kmeans and PAM with k from 2 to 20.
# Take only number of components of interest (2,5,10,20,50)
clust_res_seq_3 <- clusterMany(sim_SCE_3,ks=2:20,alphas=0.1,betas=0.8,minSizes=1,
            clusterFunction=c("kmeans","pam"), sequential=TRUE,subsample=FALSE,
                  reduceMethod="PCA",nReducedDims=c(2,5,10,20,50),verbose=TRUE)

# With ClusterExperiment object generate cluster matrix showing the clustered group names of each cell(row)
clust_res_mat_3 <- clusterMatrix(clust_res_3)
clust_res_seq_mat_3 <- clusterMatrix(clust_res_seq_3)

# For sequential clustering methods, generate an list of empty matrices   
clust_res_mat_seq_NA_3 <- matrix(nrow=nrow(clust_res_seq_mat_3),ncol=ncol(clust_res_seq_mat_3))

# Change group -1 to NA
for (b in 1:nrow(clust_res_mat_seq_NA_3)) {
    for (c in 1:ncol(clust_res_mat_seq_NA_3)) {
      clust_res_mat_seq_NA_3[b,c] <- ifelse(clust_res_seq_mat_3[b,c]=="-1",NA,clust_res_seq_mat_3[b,c])
    }
  }

# assign colnames to new list of matrices
colnames(clust_res_mat_seq_NA_3) <- colnames(clust_res_seq_mat_3)

# if a column is all NA change everything is Group1
idx <- which(colSums(is.na(clust_res_mat_seq_NA_3)) == NROW(clust_res_mat_seq_NA_3))
# result: none of the columns have all NAs
# clust_res_mat_NA_fixed <- clust_res_mat_seq_NA_3[,idx] <- 1

# divide the matrices in list to kmeans matrices and pam matrices
kmeansMat_3 <- clust_res_mat_3[,1:95]
pamMat_3 <- clust_res_mat_3[,96:190]

kmeansMat2_3 <- clust_res_mat_seq_NA_3[,1:95]
pamMat2_3 <- clust_res_mat_seq_NA_3[,96:190]

ks=2:20
```

## Clustering
### Non-sequential Kmeans
```{r}
# kmeans for each pc regardless of k
kmeans_by_pc_3 <- lapply(1:length(pc), function (i) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(kmeansMat_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
})


avg_kmean_by_pc_3 <- lapply(1:length(pc), function(i) {
      mean(unlist(kmeans_by_pc_3[[i]]))
})

avg_kmean_by_pc_3 <- unlist(avg_kmean_by_pc_3)
```

### Non-sequential PAM
```{r}
# pam for each pc regardless of k
pam_by_pc_3 <- lapply(1:length(pc), function (i) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(pamMat_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
})

avg_pam_by_pc_3 <- lapply(1:length(pc), function(i) {
      mean(unlist(pam_by_pc_3[[i]]))
    })

# Unlist for plotting
avg_pam_by_pc_3 <- unlist(avg_pam_by_pc_3)
```

### Sequantial Kmeans
```{r}
# kmeans for each pc regardless of k
kmeans_by_pc2_3 <- lapply(1:length(pc), function (i) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(kmeansMat2_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
  })

avg_kmean_by_pc2_3 <- lapply(1:length(pc), function(i) {
      mean(unlist(kmeans_by_pc2_3[[i]]))
    })

avg_kmean_by_pc2_3 <- unlist(avg_kmean_by_pc2_3)
```

### Sequential PAM
```{r}
pam_by_pc2_3 <- lapply(1:length(pc), function (i) {
    lapply(1:length(ks), function (s) {
      col_num <- 5*(s-1)+i
      comparing.Partitions(pamMat2_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
  })

avg_pam_by_pc2_3 <- lapply(1:length(pc), function(i) {
      mean(unlist(pam_by_pc2_3[[i]]))
    })

# Unlist for plotting
avg_pam_by_pc2_3 <- unlist(avg_pam_by_pc2_3)
```

## Plots for Average Rand Index(Y) for each number of Component(X) 
### Non-Sequential
```{r}
# Make a dataframe including all Indexes
avg_ari_3 <- data.frame(c(rep("kmeans",length(pc)),rep("pam",length(pc))),rep(pc,2),c(avg_kmean_by_pc_3,avg_pam_by_pc_3))

colnames(avg_ari_3) <- c("method","pc","index")

# Plot for AARI
ggplot(avg_ari_3,aes(x=pc,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

### Sequential
```{r}
# Make a dataframe including all Indexes
avg_ari2_3 <- data.frame(c(rep("kmeans",length(pc)),rep("pam",length(pc))),rep(pc,2),c(avg_kmean_by_pc2_3,avg_pam_by_pc2_3))

colnames(avg_ari2_3) <- c("method","pc","index")

# Plot for AARI
ggplot(avg_ari2_3,aes(x=pc,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

## Clustering result (for each k, PC fixed to 5)
### Non-sequential Kmeans
```{r}
## With PC=5
# kmeans by regardless of pc

kmeans_by_k3_3 <- lapply(1:length(ks), function (i) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(kmeansMat_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
  })

# Get the average for each PCs
avg_kmeans_by_k3_3 <- lapply(1:length(ks), function(i) {
  mean(unlist(kmeans_by_k3_3[[i]]))
    })

avg_kmeans_by_k3_3 <- unlist(avg_kmeans_by_k3_3)
```

### Non-sequential PAM
```{r}
# kmeans regardless of pc
pam_by_k3_3 <- lapply(1:length(ks), function (i) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(pamMat_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
})

# Get the average for each PCs
avg_pam_by_k3_3 <- lapply(1:length(ks), function(i) {
      mean(unlist(pam_by_k3_3[[i]]))
})

# Unlist for plotting
avg_pam_by_k3_3 <- unlist(avg_pam_by_k3_3)
```

### Sequential Kmeans
```{r}
kmeans_by_k4_3 <- lapply(1:length(ks), function (i) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(kmeansMat2_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
})

# Get the average for each PCs
avg_kmeans_by_k4_3 <- lapply(1:length(ks), function(i) {
  mean(unlist(kmeans_by_k4_3[[i]]))
    })

avg_kmeans_by_k4_3 <- unlist(avg_kmeans_by_k4_3)
```

### Sequential PAM
```{r}
pam_by_k4_3 <- lapply(1:length(ks), function (i) {
    lapply(1:length(pc), function (s) {
      col_num <- 5*i-3
      comparing.Partitions(pamMat2_3[,col_num],as.vector(filtered_counts3$cell_type1))
    })
})

# Get the average for each PCs
avg_pam_by_k4_3 <- lapply(1:length(ks), function(i) {
      mean(unlist(pam_by_k4_3[[i]]))
  })

# Unlist for plotting
avg_pam_by_k4_3 <- unlist(avg_pam_by_k4_3)
```

## Plots for Average Rand Index(Y) for each k (X)
### Non-Sequential
```{r}
# Make a dataframe including all Indexes
avg_ari3_3 <- data.frame(c(rep("kmeans",length(ks)),rep("pam",length(ks))),rep(ks,2),c(avg_kmeans_by_k3_3,avg_pam_by_k3_3))

colnames(avg_ari3_3) <- c("method","K","index")

# Plot for AARI
ggplot(avg_ari3_3,aes(x=K,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

### Sequential
```{r}
avg_ari4_3 <- data.frame(c(rep("kmeans",length(ks)),rep("pam",length(ks))),rep(ks,2),c(avg_kmeans_by_k4_3,avg_pam_by_k4_3))

colnames(avg_ari4_3) <- c("method","K","index")

# Plot for AARI
ggplot(avg_ari4_3,aes(x=K,y=index,group=method,col=method)) +
  geom_smooth(stat = "identity", position = "identity") +
  theme_bw()
```

```{r}
save(avg_ari_3,avg_ari2_3,avg_ari3_3,avg_ari4_3, file="/Users/jieun/Desktop/CapstoneProject/FinalScripts/data3results.Rdata")

```

 